### 2016-08-28 Chapters/ContinuumSolvationModels.tex
Save work on Chapter 1

\begin{itemize}
    \item why and how continuum solvation models
    \item derivation of \acs{IEF}-\acs{PCM}, importance of Green's
      functions.
    \item \ac{BEM} and wavelet \acs{BEM} I think it's necessary to
      mention also ddCOSMO and ddPCM as alternative strategies
      ~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
      Lipparini2014-fq, Stamm2016-fv}
    \item variational formulation (no derivations, just mention the
      relevant literature)
\end{itemize}

@@

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.65\textwidth]{IEF}
  \caption{Vacuum Cavity in dielectric with permittivity $\diel_\mathrm{e}>1$.
  Source charge density $\rho_\mathrm{i}$ inside the cavity
  ($\diel_\mathrm{i}=1$). \textcolor{red}{Regular \acs{IEF}}}
  \label{fig:case0}
\end{figure}

@@

Approximability, conformity, consistency and asymptotic consistency
of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
TesiFilippo}
Additional cavity generators
TsLess~\autocite{Pomelli2004-lb}
DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc}
FIXPVA~\autocite{Su2009-em}
Other discretization schemes
CSC~\autocite{York1999-xy, Scalmani2010-tw}
SWIG~\autocite{Lange2010-jp, Lange2010-qo, Lange2011-eu}

>>>>

Developments of the \ac{PCM}, a continuum solvation model, are at the
heart of this thesis.
This Chapter is a brief exposition of the major points of the \acs{PCM}
we have worked upon in the thesis.
Section \ref{sec:IEF} will present some of the mathematical details of the
formulation of the model. I will introduce the concept of \emph{weak}
formulations of \acp{PDE} in Section \ref{sec:weak} and eventually show their
use in the formulation of quantum/classical polarizable Hamiltonians in
Section \ref{sec:variational}.
Section \ref{sec:BEM} will be devoted to the numerical strategies used
in solving the \acs{PCM} equation.

Continuum models have a venerable history in quantum chemistry:
\citeauthor{Onsager1936-wf}'s model\autocite{Onsager1936-wf} appeared in
1936 and the first form of the \acs{PCM} entered the stage in
1981.~\autocite{Miertus1981-mm}
Section \ref{sec:CSM-why-how} will offer a heuristic "derivation" of
continuum models.
While it cannot be thought as formally rigorous, it highlights the major
physical insights that have informed the creation of continuum models.
Still nowadays much research activity is expended on the
\acs{PCM}.~\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
I will point out throughout the Chapter to developments outside our
group that are particularly interesting.

@@

\section{Continuum Solvation Models: Why and How}\label{sec:CSM-why-how}

The basic idea of implicit models is that the environment can
be replaced by structureless continuum. The continuum will interact
\emph{classically} with the explicit part of the multiscale model, an
interaction mediated by some macroscopic property of the bulk material
the continuum represents in the model.
Thus, continuum models bypass both problems plaguing explicit models we
mentioned in the Introduction, namely:
\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
\item the statistical averaging of environment configurations and
\item the \acs{MM} region electrostatics cutoff choice.
\end{enumerate*}

Following \citeauthor{Tomasi2007-es}, we introduce the complete system
Hamiltonian in the \acs{BO} approximation:~\autocite{Tomasi2004-dc,
Tomasi2007-es}
\begin{equation}\label{eq:solution-ham}
 H(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E}) =
  H_\mathrm{S}(\vect{r}_\mathrm{S}) +  H_\mathrm{E}(\vect{r}_\mathrm{E})
+ H_\mathrm{SE}(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E})
\end{equation}
The Hamiltonian features solute terms, marked by the S subscript,
environment terms, marked by the E subscript, and interaction terms.
The coordinates $(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E})$ refer to
\emph{both} nuclei and electrons.
The interaction term is given by the usual Coulomb electrostatic
Hamiltonian.
\todo[inline]{What is a nonpolarizable model? What's a polarizable
model? HAS TO BE MENTIONED HERE!}
\todo[inline]{What is the nature of the interaction? Mention that we
only include polarization, but in a mutual manner.}

One can replace the pure environment and interaction terms with their
classical counterparts and obtain the quantum/classical, possibly
\emph{polarizable}, Hamiltonian for an explicit \acs{QM}/\acs{MM} model.
This brings about the first important point: the need for statistical
averaging.
Whenever a large number of degrees of freedom is involved, one can
access macroscopic observables of the system by the appropriate
averaging of the microscopically detailed motion over phase space
trajectories or on the appropriate statistical
ensemble.~\autocite{Hill1960-ql}
The need for ensemble averages leads us to the following observations:
\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
 \item
   the need for \emph{macroscopic} parameters, absent from the
   microscopic Hamiltonian Eq.~\eqref{eq:solution-ham}, in carrying out
   statistical simulations.
 \item
   Chosen a thermodynamic ensemble, the basic energetic quantity is
   accordingly determined. For example, the Gibbs free energy  $G$ is
   intrinsically related to the $(NpT)$ ensemble.
 \item
   Even atomistic simulations will lead to results that are essentially
   averages. The discrete picture of the system has been replaced by
   some continuous distribution function.
\end{enumerate*}

These observations strongly suggest that one can perform the averaging
step \emph{before} embarking into the solution of the quantum mechanical
problem and replace the full Hamiltonian with and \emph{effective}
Hamiltonian in which the environment degrees of freedom are properly
averaged.~\autocite{Angyan1992-vo, Tapia1992-pu}
The solute-environment interaction term would then be replaced by
a term only depending on the solute degrees of freedom, represented by
the averaged, continuous response functions of the solvent.

\todo[inline]{Introduction of the cavity. PCM as a problem in classical
electrostatics}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.65\textwidth]{IEF}
  \caption{
  The physical setting of the \ac{PCM}. The molecular solute,
  represented by its, possibly quantum mechanical, charge density
  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
  2-manifold and is assumed to be $\mathcal{C}^1.
  We assume the material properties of the cavity to be those of vacuum,
  hence characterized by the Green's function
  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
  The cavity is carved out of an infinite, structureless, continuum
  characterized by its Green's function $\Ge$ and fully covering the
  external subdomain $\Omegae$.}
  \label{fig:IEF}
\end{figure}

@@

@@

Approximability, conformity, consistency and asymptotic consistency
of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
TesiFilippo}
Additional cavity generators
TsLess~\autocite{Pomelli2004-lb}
DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc}
FIXPVA~\autocite{Su2009-em}
Other discretization schemes
CSC~\autocite{York1999-xy, Scalmani2010-tw}
SWIG~\autocite{Lange2010-jp, Lange2010-qo, Lange2011-eu}
ddCOSMO and ddPCM~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
Lipparini2014-fq, Stamm2016-fv}
